### 2023-11-30 0.Intro/resumo.tex
Changed a bunch of files
Added almost all symbols

\begin{abstract}

  Declaration
  I declare that this document is an original work of my own authorship and that it fulfills all
  the requirements of the Code of Conduct and Good Practices of the Universidade de
  Lisboa.
  In this work we propose an analog structure for memristor based recurrent modules targeting neural computing. The system is fully analog and implements a working \ac{LSTM} circuit block and a work in progress \ac{GRU} circuit block. Both of those blocks contain memristors to be used as weights in a analog \ac{VMM} capable circuit. These circuit blocks allow to run very fast computation of \aclp{RNN} of any size, in a relatively small integrated circuit. As part of the \ac{LSTM} and \ac{GRU} blocks, an analog activation function circuit was designed. This specific circuit is capable of reproducing sigmoid and \ac{tanh} like functions, with similar shapes and the same output ranges. The work also include the implementation of a memory cell used to store an analog value for a short period of time. The \ac{LSTM} block can be serialized or not with the ability to choose the level of serialization. Serializing the system allows to save onChip area at the cost of execution time. To the author's knwoledge this is the first analog implementation of the behavior of \acs{C. elegans} using the \ac{LSTM} block. Such an analog system provides ground for real time implementation of nervous systems.

\end{abstract}

>>>>

\begin{resumo}

  Neste trabalho, propomos uma estrutura anal\'{o}gica para m\'{o}dulos recorrentes baseados em memristors para computa\c{c}\~{a}o neurom\'{o}rfica. O sistema \'{e} totalmente anal\'{o}gico e implementa um de circuito \acs{LSTM} funcional bem como um circuito GRU, ainda em desenvolvimento. Ambos os blocos cont\^{e}m memristors para serem usados como pesos em um circuito VMM anal\'{o}gico capaz. Esses blocos permitem a execu\c{c}\~{a}o de c\'{a}lculos necess\'{a}rios para Redes Neurais Recorrentes de dimens\~{a}o arbitr\'{a}ria, e de forma rapida e eficiente, em circuito integrado. Como parte dos blocos \ac{LSTM} e \ac{GRU}, foi projetado um circuito anal\'{o}gico que implementa fun\c{c}\~{o}es de ativa\c{c}\~{a}o. Este circuito espec\'{i}fico \'{e} capaz de reproduzir fun\c{c}\~{o}es semelhantes \`{a}s sigmoid e tanh, com formas e intervalos de valor similares. O trabalho tamb\'{e}m inclui a implementa\c{c}\~{a}o de uma c\'{e}lula de mem\'{o}ria usada para armazenar um valor anal\'{o}gico por um curto per\'{i}odo de tempo. Os c\'{a}lculos do bloco LSTM podem ser executado em s\'{e}rie ou paralelo, com um grau de serializa\c{c}\~{a}o arbitr\'{a}rio. A serializa\c{c}\~{a}o do sistema permite economizar \'{a}rea no chip, a custo do tempo de execu\c{c}\~{a}o. Adicionalmente, at\'{e} \`{a} data de publica\c{c}\~{a}o esta \'{e} a primeira implementa\c{c}\~{a}o anal\'{o}gica do comportamento do \ac{C. elegans} usando o bloco \ac{LSTM} em circuito integrado. Um sistema anal\'{o}gico deste tipo estabelece a base para a implementa\c{c}\~{a}o em tempo real de sistemas nervosos.

\end{resumo}

### 2023-11-30 design/crossbar.tex
Added almost all symbols

This design is the one that is used in \cite{doubleMem}. Let's assume that a given memristor has a resistance range of $R\in[R_{min},R_{max}]$, that means its conductance range is $ G \in [ G_{min}, G_{max}]$ (with $ G_{min}= \frac{1}{R_{max}}$ and $ G_{max}= \frac{1}{R_{min}}$). This design works using two \ac{opAmp} connected to $V_{cm}$ with the positive pin and the negative pin to the output of the crossbar array. \Cref{eq:doubleMem0,eq:doubleMem1,eq:doubleMem2} are describing how this architecture works. A simplified version of the double memristors per synapse circuit is also available in \cref{circt:doubleMem}.

@@

Using $x_k$ as the voltage for the input line $k$. The highest \ac{opAmp} is identified as $opAmp_0$ and the lowest $opAmp_1$.

For the sake of simplicity, the following equations considers the ground to be $V_{cm}$.

\begin{equation}
  \label{eq:doubleMem0}
  V_{opAmp_0}=-R_r\cdot i_+
\end{equation}
\begin{equation}
  \label{eq:doubleMem1}
  i_{R_f}=i_-+\frac{V_{opAmp_0}}{R_r}=i_--i_+
\end{equation}
\begin{equation}
  \label{eq:doubleMem2}
  V_{opAmp_1}=y_0=-R_f\cdot(i_--i_+)=R_f\cdot(i_+-i_-)=R_f\cdot\sum_{k=0}^n( G_{k+}- G_{k-})\cdot x_k
\end{equation}
With $i_+=\sum_{k=0}^n G_{k+}\cdot x_k$ and $i_-=\sum_{k=0}^n G_{k-}\cdot x_k$.

\begin{table}[H]
  \centering
  \begin{tabular}{|c|c|c|}
    \cline{2-3}
    \rowcolor{gray}
    \multicolumn{1}{c|}{\cellcolor[HTML]{FFFFFF}} & Two memristors per synapse & One memristor per synapse \\
    \hline
    Maximum weight & $ G_{max}- G_{min}$ & $ G_{max} -\overline{ G}$\\
    \hline
    Minimum weight & $ G_{min}- G_{max}$ & $ G_{min} -\overline{ G}$\\
    \hline
    Range & $2\cdot( G_{max}- G_{min})$&$ G_{max}- G_{min}$\\
    \hline
  \end{tabular}
  \caption{Synaptic weights precision (extracted from \cite{doubleMem})}
  \label{tab:synapses}
\end{table}

@@

\begin{equation}
  A_{xbar}(n_i,n_o,n_s)=
  \begin{cases}
    2\cdot n_i\cdot n_o \cdot A_{memristor}+n_o\cdot(2\cdot[A_{opAmp}+A_{R_r}]+A_{R_f})& n_s=1\\
    \begin{array}{2}
      2\cdot n_i\cdot n_o \cdot n_s\cdot A_{memristor}+2\cdot A_{CMOS}\cdot n_o\cdot n_s\\
      +n_o\cdot(2\cdot[A_{opAmp}+A_{R_r}]+A_{R_f})
    \end{array}
    & \forall n_s>1\\
  \end{cases}
\end{equation}

With $A_{CMOS}=w\cdot l$.

@@

\begin{equation}\label{eq:memThres}
  |V_{mem}|= |x_i-V_{cm}|\le V_{read}
\end{equation}

Where $V_{mem}$ is the terminal voltage of a memristor.

>>>>

This design is the one that is used in \cite{doubleMem}. Let's assume that a given memristor has a resistance range of $\symR\in[\symR_{min},\symR_{max}]$, that means its conductance range is $ \symG \in [ \symG_{min}, \symG_{max}]$ (with $ \symG_{min}= \frac{1}{\symR_{max}}$ and $ \symG_{max}= \frac{1}{\symR_{min}}$). This design works using two \ac{opAmp} connected to $\symv_{cm}$ with the positive pin and the negative pin to the output of the crossbar array. \Cref{eq:doubleMem0,eq:doubleMem1,eq:doubleMem2} are describing how this architecture works. A simplified version of the double memristors per synapse circuit is also available in \cref{circt:doubleMem}.

@@

Using $\symv_k$ as the voltage for the input line $k$. The highest \ac{opAmp} is identified as $opAmp_0$ and the lowest $opAmp_1$.

For the sake of simplicity, the following equations considers the ground to be $\symv_{cm}$.

\begin{equation}
  \label{eq:doubleMem0}
  \symv_{opAmp_0}=-\symR_r\cdot \symi_+
\end{equation}
\begin{equation}
  \label{eq:doubleMem1}
  \symi_{\symR_f}=\symi_-+\frac{\symv_{opAmp_0}}{\symR_r}=\symi_--\symi_+
\end{equation}
\begin{equation}
  \label{eq:doubleMem2}
  \symv_{opAmp_1}=\symv_{out}=-\symR_f\cdot(\symi_--\symi_+)=\symR_f\cdot(\symi_+-\symi_-)=\symR_f\cdot\sum_{k=0}^n( \symG_{k+}- \symG_{k-})\cdot \symv_k
\end{equation}
With $\symi_+=\sum_{k=0}^n \symG_{k+}\cdot \symv_k$ and $\symi_-=\sum_{k=0}^n \symG_{k-}\cdot \symv_k$.

\begin{table}[H]
  \centering
  \begin{tabular}{|c|c|c|}
    \cline{2-3}
    \rowcolor{gray}
    \multicolumn{1}{c|}{\cellcolor[HTML]{FFFFFF}} & Two memristors per synapse & One memristor per synapse \\
    \hline
    Maximum weight & $ \symG_{max}- \symG_{min}$ & $ \symG_{max} -\overline{ \symG}$\\
    \hline
    Minimum weight & $ \symG_{min}- \symG_{max}$ & $ \symG_{min} -\overline{ \symG}$\\
    \hline
    Range & $2\cdot( \symG_{max}- \symG_{min})$&$ \symG_{max}- \symG_{min}$\\
    \hline
  \end{tabular}
  \caption{Synaptic weights precision (extracted from \cite{doubleMem})}
  \label{tab:synapses}
\end{table}

@@

\begin{equation}
  \symA_{xbar}(n_i,n_o,n_s)=
  \begin{cases}
    2\cdot n_i\cdot n_o \cdot \symA_{memristor}+n_o\cdot(2\cdot[\symA_{opAmp}+\symA_{R_r}]+\symA_{R_f})& n_s=1\\
    \begin{array}{2}
      2\cdot n_i\cdot n_o \cdot n_s\cdot \symA_{memristor}+2\cdot \symA_{CMOS}\cdot n_o\cdot n_s\\
      +n_o\cdot(2\cdot[\symA_{opAmp}+\symA_{R_r}]+\symA_{R_f})
    \end{array}
    & \forall n_s>1\\
  \end{cases}
\end{equation}

With $\symA_{CMOS}=w\cdot l$.

@@

\begin{equation}\label{eq:memThres}
  |\symv_{mem}|= |\symv_i-\symv_{cm}|\le \symv_{read}
\end{equation}

Where $\symv_{mem}$ is the terminal voltage of a memristor.

### 2023-11-30 design/genwei.tex
Added almost all symbols

The weights need to be contrained during the training to make sure that the voltage of the analog circuit stays within its operating voltage range ($[0,V_{dd}]$). To avoid any unwanted behavior when the system supposedly goes out of range, the weights in the \ac{LSTM} are contrained. This constrain is not fixed and depends on the architecture it is generated with.

The worst case scenario for an \ac{LSTM} is when every value reaches the voltage threshold ($V_{read}$) (\cref{sec:xbarCircuit}) and every weight is maximized ($w_{max}$). When that is the case the output of the \acp{VMM} is found in \cref{eq:weiCons}.

\begin{equation}\label{eq:weiCons}
  V_{read}\cdot w_{max} \cdot(n_i+n_h)+w_{max}+V_{cm}= V_{dd}
\end{equation}

We can then determine the maximum acceptable value for the weights ($w_{max}$), this value has thus been determined to be the one in \cref{eq:weiConsRes}.

\begin{equation}\label{eq:weiConsRes}
  w_{max}=\frac{V_{dd}-V_{cm}}{V_{read}\cdot(n_i+n_h)+1}
\end{equation}

The analog system being completly centered and symetrical around $V_{cm}$ the value for the minimum weight is the opposite to the maximum one (\cref{eq:weiConsRes1})

\begin{equation}\label{eq:weiConsRes1}
  w_{min}=-w_{max}=-\frac{V_{dd}-V_{cm}}{V_{read}\cdot(n_i+n_h)+1}
\end{equation}

>>>>

The weights need to be contrained during the training to make sure that the voltage of the analog circuit stays within its operating voltage range ($[0,\symv_{dd}]$). To avoid any unwanted behavior when the system supposedly goes out of range, the weights in the \ac{LSTM} are contrained. This constrain is not fixed and depends on the architecture it is generated with.

The worst case scenario for an \ac{LSTM} is when every value reaches the voltage threshold ($\symv_{read}$) (\cref{sec:xbarCircuit}) and every weight is maximized ($\symw_{max}$). When that is the case the output of the \acp{VMM} is found in \cref{eq:weiCons}.

\begin{equation}\label{eq:weiCons}
  \symv_{read}\cdot \symw_{max} \cdot(n_i+n_h)+\symw_{max}+\symv_{cm}= \symv_{dd}
\end{equation}

We can then determine the maximum acceptable value for the weights ($\symw_{max}$), this value has thus been determined to be the one in \cref{eq:weiConsRes}.

\begin{equation}\label{eq:weiConsRes}
  \symw_{max}=\frac{\symv_{dd}-\symv_{cm}}{\symv_{read}\cdot(n_i+n_h)+1}
\end{equation}

The analog system being completly centered and symetrical around $\symv_{cm}$ the value for the minimum weight is the opposite to the maximum one (\cref{eq:weiConsRes1})

\begin{equation}\label{eq:weiConsRes1}
  \symw_{min}=-\symw_{max}=-\frac{\symv_{dd}-\symv_{cm}}{\symv_{read}\cdot(n_i+n_h)+1}
\end{equation}

### 2023-11-30 design/gru.tex
Changed a bunch of files
Added almost all symbols

\begin{itemize}
  \item $\overrightarrow{x_t}$ : This is the input vector for the \ac{LSTM} circuit at time $t$. It has a size of $n_i$.
  \item $\overrightarrow{h_t}$ : This is the hidden layer vector for the feedback connections, it is defined as $\overrightarrow{h_{t}}=(h_{t,i}) \forall i\in [\![0,n_h-1]\!]$.
  \item $\overrightarrow{z_t}$ is the input of the crossbar but not the input of the \ac{LSTM}. This vector is there to lighten the informations on the schematic (\cref{circt:lstm}). $\overrightarrow{z_t}$ is defined by $\overrightarrow{z_t}=(\overrightarrow{x_t},\overrightarrow{h_{t-1}},\overrightarrow{b})$.
  \item $e_{in}$ and $e_{out}$ are the flags used to enable the hidden state values to go to the input (feedback connection) or to the output of the circuit.
  \item $e_{in,end}$ is the second half of the $e_{in}$ flag.
  \item $e_{next}$ is the enable flag on in between two time steps.
  \item $R_{amp0}$ and $R_{amp1}$ are the two resistances used to amplify the output voltage of the voltage multipliers. The ratio of the resistances value must of $\frac{R_{amp1}}{R_{amp0}}=10$. The resistances must stay arround the values of the resistances used around the circuit, especially those of the memristors.
  \item $V_{inv}$ is the voltage used to perform the $1-x$ operation.
\end{itemize}

@@

\subsubsection{The $1-x$ operation}

@@

The solution found to do compute the operation consist of using an \ac{opAmp} as an inverter around $V_{inv}$, the formula is available in \cref{eq:1minus1}.
\begin{equation}\label{eq:1minus1}
  f_v(v)=-(v-V_{inv})+V_{inv}=2\cdot V_{inv} -v
\end{equation}

@@

\begin{equation}\label{eq:1minus2}
  f_v(v)=(1-V_{cm})-(v-V_{cm})+V_{cm}= 0.1-v+2\cdot V_{cm}
\end{equation}

It can thus be found that the value for $V_{inv}$ is $V_{inv}=V_{cm}+0.05$

@@

\begin{equation}\label{eq:serProb}
  \overrightarrow{\hat{h_t}}=tanh(\overrightarrow{x_t}\cdot w_{hx}+(\overrightarrow{r_t}\odot\overrightarrow{h_{t-1}}) \cdot w_{hh} + \overrightarrow{b_h})
\end{equation}

@@

\begin{equation}
  \begin{array}{3}
    A_{gru}(n_i,n_h)=4\cdot A_{xbar}(n_i,n_h,1)\\+ \frac{n_h}{n_s}\cdot (5\cdot A_{af}+3\cdot A_{voltMult}+3\cdot A_{R_{amp0}}+2\cdot A_{R_{amp1}}+2\cdot A_{opAmp})\\+5\cdot n_h\cdot A_{memcell}
  \end{array}
\end{equation}

>>>>

\begin{itemize}
  \item $\overrightarrow{x_t}$ : This is the input vector for the \ac{LSTM} circuit at time $t$. It has a size of $n_i$.
  \item $\overrightarrow{h_t}$ : This is the hidden layer vector for the feedback connections, it is defined as $\overrightarrow{h_{t}}=(h_{t,i}) \forall i\in [\![0,n_h-1]\!]$.
  \item $\overrightarrow{z_t}$ is the input of the crossbar but not the input of the \ac{LSTM}. This vector is there to lighten the informations on the schematic (\cref{circt:lstm}). $\overrightarrow{z_t}$ is defined by $\overrightarrow{z_t}=(\overrightarrow{x_t},\overrightarrow{h_{t-1}},\overrightarrow{b})$.
  \item $e_{in}$ and $e_{out}$ are the flags used to enable the hidden state values to go to the input (feedback connection) or to the output of the circuit.
  \item $e_{in,end}$ is the second half of the $e_{in}$ flag.
  \item $e_{next}$ is the enable flag on in between two time steps.
  \item $\symR_{amp0}$ and $\symR_{amp1}$ are the two resistances used to amplify the output voltage of the voltage multipliers. The ratio of the resistances value must of $\frac{\symR_{amp1}}{\symR_{amp0}}=10$. The resistances must stay arround the values of the resistances used around the circuit, especially those of the memristors.
  \item $\symv_{inv}$ is the voltage used to perform the $1-x$ operation.
\end{itemize}

@@

\subsection{The $1-x$ operation}

@@

The solution found to do compute the operation consist of using an \ac{opAmp} as an inverter around $V_{inv}$, the formula is available in \cref{eq:1minus1}.
\begin{equation}\label{eq:1minus1}
  f_\symv(\symv)=-(\symv-\symv_{inv})+\symv_{inv}=2\cdot \symv_{inv} -\symv
\end{equation}

@@

\begin{equation}\label{eq:1minus2}
  f_\symv(\symv)=(1-\symv_{cm})-(\symv-\symv_{cm})+\symv_{cm}= 0.1-\symv+2\cdot \symv_{cm}
\end{equation}

It can thus be found that the value for $\symv_{inv}$ is $\symv_{inv}=\symv_{cm}+0.05$

@@

\begin{equation}\label{eq:serProb}
  \symvhh_t=tanh(\symvx_t\cdot \symmw_{hx}+(\symvr_t\odot\symvh_{t-1}) \cdot \symmw_{hh} + \symvb_h)
\end{equation}

@@

\begin{equation}
  \begin{array}{3}
    \symA_{gru}(n_i,n_h)=4\cdot \symA_{xbar}(n_i,n_h,1)\\+ \frac{n_h}{n_s}\cdot (5\cdot \symA_{af}+3\cdot \symA_{voltMult}+3\cdot \symA_{\symR_{amp0}}+2\cdot \symA_{\symR_{amp1}}+2\cdot \symA_{opAmp})\\+5\cdot n_h\cdot \symA_{memcell}
  \end{array}
\end{equation}

### 2023-11-30 design/lstm.tex
Worked on paper+corrected errors
Added almost all symbols

First of all, the different vectors and variables present in the schematic have to be described :

\begin{itemize}
  \item $\overrightarrow{x_t}$ : This is the input vector for the \ac{LSTM} circuit at time $t$. It has a size of $n_i$.
  \item $\overrightarrow{h_t}$ : This is the hidden layer vector for the feedback connections, it is defined as $\overrightarrow{h_t}=(\overrightarrow{h_{t,i}}) \forall i\in [\![0,n_o-1]\!]$, with $\overrightarrow{h_{t,i}}=(h_{t,i,j}) \forall j\in [\![0,n_s-1]\!]$.
  \item $\overrightarrow{z_t}$ is the input of the crossbar but not the input of the \ac{LSTM}. This vector is there to lighten the informations on the schematic (\cref{circt:lstm}). $\overrightarrow{z_t}$ is defined by $\overrightarrow{z_t}=(\overrightarrow{x_t},\overrightarrow{h_{t-1}},1)$.
  \item $e_{j,0}$ and $e_{j,1}$ are two enable flags that respectively represent the first and second half of $e_j$.
  \item $e_{in}$ and $e_{out}$ are the flags used to enable the hidden state values to go to the input (feedback connection) or to the output of the circuit.
  \item $e_{next}$ is the enable flag on in between two time steps.
  \item $R_{amp0}$ and $R_{amp1}$ are the two resistances used to amplify the output voltage of the voltage multipliers. The ratio of the resistances value must of $\frac{R_{amp1}}{R_{amp0}}=10$. The resistances must stay arround the values of the resistances used around the circuit, especially those of the memristors.
\end{itemize}

@@

\begin{equation}
  \begin{array}{3}
    A_{lstm}(n_i,n_h,n_s)=4\cdot A_{xbar}(n_i,\frac{n_h}{n_s},n_s)\\+ \frac{n_h}{n_s}\cdot (5\cdot A_{af}+3\cdot A_{voltMult}+3\cdot A_{R_{amp0}}+2\cdot A_{R_{amp1}}+2\cdot A_{opAmp})\\+5\cdot n_h\cdot A_{memcell}
  \end{array}
\end{equation}

>>>>

The different vectors and variables present in the schematic have to be described :

\begin{itemize}
  \item $\overrightarrow{x_t}$ : This is the input vector for the \ac{LSTM} circuit at time $t$. It has a size of $n_i$.
  \item $\overrightarrow{h_t}$ : This is the hidden layer vector for the feedback connections, it is defined as $\overrightarrow{h_t}=(\overrightarrow{h_{t,i}}) \forall i\in [\![0,n_o-1]\!]$, with $\overrightarrow{h_{t,i}}=(h_{t,i,j}) \forall j\in [\![0,n_s-1]\!]$.
  \item $\overrightarrow{z_t}$ is the input of the crossbar but not the input of the \ac{LSTM}. This vector is there to lighten the informations on the schematic (\cref{circt:lstm}). $\overrightarrow{z_t}$ is defined by $\overrightarrow{z_t}=(\overrightarrow{x_t},\overrightarrow{h_{t-1}},1)$.
  \item $e_{j,0}$ and $e_{j,1}$ are two enable flags that respectively represent the first and second half of $e_j$.
  \item $e_{in}$ and $e_{out}$ are the flags used to enable the hidden state values to go to the input (feedback connection) or to the output of the circuit.
  \item $e_{next}$ is the enable flag on in between two time steps.
  \item $\symR_{amp0}$ and $\symR_{amp1}$ are the two resistances used to amplify the output voltage of the voltage multipliers. The ratio of the resistances value must of $\frac{\symR_{amp1}}{\symR_{amp0}}=10$. The resistances must stay arround the values of the resistances used around the circuit, especially those of the memristors.
\end{itemize}

@@

\begin{equation}
  \begin{array}{3}
    \symA_{lstm}(n_i,n_h,n_s)=4\cdot \symA_{xbar}(n_i,\frac{n_h}{n_s},n_s)\\+ \frac{n_h}{n_s}\cdot (5\cdot \symA_{af}+3\cdot \symA_{voltMult}+3\cdot \symA_{\symR_{amp0}}+2\cdot \symA_{\symR_{amp1}}+2\cdot \symA_{opAmp})\\+5\cdot n_h\cdot \symA_{memcell}
  \end{array}
\end{equation}

### 2023-11-30 design/main.tex
Correcting inappropriate raccourcis and correcting design
Added almost all symbols

The system will be working with a $V_{dd}$ of $1.8V$. Such a value was chosen because this is a low power system.

The way values are encoded in the analog system will be descibed here as it serves for the entire thesis.
In order for the system to support negative numbers we're going to use a $V_{cm}$ set to $\frac{V_{dd}}{2}$. That means that $V_{cm}=\frac{V_{dd}}{2}=0.9V$. This $V_{cm}$ will then describe a zero. A step of one was chosen to be $0.1V$ in the analog circuit.
\Cref{tab:valConv} shows the conversions from a real number to it's voltage equivalent.

\begin{table}[H]
  \centering
  \begin{tabular}{|c|c|}
    \hline
    \rowcolor{gray}
    Real value & Voltage \\
    \hline
    $0$ & $0.9V$ \\
    \hline
    $1$ & $1.0V$ \\
    \hline
    $x$ & $V(x)=\frac{x}{10}+V_{cm}$\\
    \hline
    $real(v)=(v-0.9)\cdot 10$ & $v$\\
    \hline
  \end{tabular}
  \caption{Real/Voltage Conversion Table.}
  \label{tab:valConv}
\end{table}

Since the system cannot reach voltage outside of the operating range with the intended behavior, the voltage is then restricted to $V\in [0,1.8]$. This means that the range of real value that the systems can handle is $x\in [-9,9]$.

>>>>

The system will be working with a $\symv_{dd}$ of $1.8V$. Such a value was chosen because this is a low power system.

The way values are encoded in the analog system will be descibed here as it serves for the entire thesis.
In order for the system to support negative numbers we're going to use a $\symv_{cm}$ set to $\frac{\symv_{dd}}{2}$. That means that $\symv_{cm}=\frac{\symv_{dd}}{2}=0.9V$. This $\symv_{cm}$ will then describe a zero. A step of one was chosen to be $0.1V$ in the analog circuit.
\Cref{tab:valConv} shows the conversions from a real number to its voltage equivalent.

\begin{table}[H]
  \centering
  \begin{tabular}{|c|c|}
    \hline
    \rowcolor{gray}
    Real value & Voltage \\
    \hline
    $0$ & $0.9V$ \\
    \hline
    $1$ & $1.0V$ \\
    \hline
    $x$ & $\symv(x)=\frac{x}{10}+\symv_{cm}$\\
    \hline
    $real(\symv)=(\symv-\symv_{cm})\cdot 10$ & \symv\\
    \hline
  \end{tabular}
  \caption{Real/Voltage Conversion Table.}
  \label{tab:valConv}
\end{table}

Since the system cannot reach voltage outside of the operating range with the intended behavior, the voltage is then restricted to $\symv\in [0,1.8]$. This means that the range of real value that the systems can handle is $x\in [-9,9]$.

### 2023-11-30 design/memcell.tex
Added almost all symbols

\begin{equation}
  A_{memcell}=8\cdot A_{CMOS}+2\cdot A_{inv}+A_{opAmp}+A_{cap}
\end{equation}

With $A_{CMOS} = w\cdot l$

@@

This circuit is pretty straight forward, when the input enable pin is high (up to $V_{dd}$), then the first switch is open and the capacitor is being charged, if it is low, then the switch is closed and the capacitor keeps the voltage and thus the analog data. When the output enable is high, the buffer (\cref{apsec:buffer}) forwards the voltage stored in the capactor to the output pin without emptying the capacitor.

>>>>

\begin{equation}
  \symA_{memcell}=8\cdot \symA_{CMOS}+2\cdot \symA_{inv}+\symA_{opAmp}+\symA_{cap}
\end{equation}

With $\symA_{CMOS} = w\cdot l$

@@

This circuit is pretty straight forward, when the input enable pin is high (up to $\symv_{dd}$), then the first switch is open and the capacitor is being charged, if it is low, then the switch is closed and the capacitor keeps the voltage and thus the analog data. When the output enable is high, the buffer (\cref{apsec:buffer}) forwards the voltage stored in the capactor to the output pin without emptying the capacitor.

### 2023-11-30 design/models.tex
Added almost all symbols

\begin{equation}
  \label{eq:opAmp}
  V_{out}=\mu \cdot (V_+-V_-)
\end{equation}

@@

However this part is tricky. Indeed, because of the voltage conversion (\cref{tab:valConv}), the equation needs to take the conversion into consideration. The final equation needs to be \cref{eq:finalVoltMult} in order to have $1\cdot 1=1$ ($V_{in_1}\cdot V_{in_2}=1V$, with $V_{in_1}=V_{in_2}=1V$)

\begin{equation}\label{eq:finalVoltMult}
  V_{out}=10\cdot(V_{in_1}-V_{cm})\cdot (V_{in_2}-V_{cm}) + V_{cm}
\end{equation}

@@

\begin{equation}\label{eq:voltMult}
  V_{voltMult}=-(V_{in_1}-V_{cm})\cdot (V_{in_2}-V_{cm}) + V_{cm}
\end{equation}
\begin{equation}\label{eq:invAmp}
  V_{out}=-(V_{voltMult}-V_{cm})\cdot10+V_{cm}
\end{equation}

Where $V_{out}$ is the output voltage the inverting amplifier (\cref{apsec:sumInv}), $V_{voltMult}$ is the out voltage of the voltage multiplier itself and $V_{in_1}$ and $V_{in_2}$ are the input voltages.

>>>>

\begin{equation}
  \label{eq:opAmp}
  \symv_{out}=\mu \cdot (\symv_+-\symv_-)
\end{equation}

@@

However this part is tricky. Indeed, because of the voltage conversion (\cref{tab:valConv}), the equation needs to take the conversion into consideration. The final equation needs to be \cref{eq:finalVoltMult} in order to have $1\cdot 1=1$ ($\symv_{in_1}\cdot \symv_{in_2}=1V$, with $\symv_{in_1}=\symv_{in_2}=1V$)

\begin{equation}\label{eq:finalVoltMult}
  \symv_{out}=10\cdot(\symv_{in_1}-\symv_{cm})\cdot (\symv_{in_2}-\symv_{cm}) + \symv_{cm}
\end{equation}

@@

\begin{equation}\label{eq:voltMult}
  \symv_{voltMult}=-(\symv_{in_1}-\symv_{cm})\cdot (\symv_{in_2}-\symv_{cm}) + \symv_{cm}
\end{equation}
\begin{equation}\label{eq:invAmp}
  \symv_{out}=-(\symv_{voltMult}-\symv_{cm})\cdot10+\symv_{cm}
\end{equation}

Where $\symv_{out}$ is the output voltage the inverting amplifier (\cref{apsec:sumInv}), $\symv_{voltMult}$ is the out voltage of the voltage multiplier itself and $\symv_{in_1}$ and $\symv_{in_2}$ are the input voltages.

### 2023-11-30 design/run.tex
Added almost all symbols

In order to run the simulation some variables need to be set. The time step ($T$) value was chosen to be $T=8\mu s$, the pause between two \ac{LSTM} steps is set to be $\frac{T}{8}=1\mu s$. The value for the maximum and minimum resistances value of the memristor used ($R_{max}$ and $R_{min}$, respectively) are set to be in acordance with the \textbf{KNOWN}'s memristors. The resistance values used for the simulation are thus $R_{max}=10^6 \Omega=1M\Omega$ and $R_{min}=10^4 \Omega = 10k\Omega$.

>>>>

In order to run the simulation some variables need to be set. The time step ($T$) value was chosen to be $T=8\mu s$, the pause between two \ac{LSTM} steps is set to be $\frac{T}{8}=1\mu s$. The value for the maximum and minimum resistances value of the memristor used ($\symR_{max}$ and $\symR_{min}$, respectively) are set to be in acordance with the \textbf{KNOWM}\textregistered{}'s memristors \cite{Knowm}. The resistance values used for the simulation are thus $\symR_{max}=10^6 \Omega=1M\Omega$ and $\symR_{min}=10^4 \Omega = 10k\Omega$.

### 2023-11-30 design/wei2res.tex
Worked on appendix
corrected a few things
Added almost all symbols

\begin{equation}
  \label{eq:wei2res0}
  w=R_f\cdot(\sigma_+-\sigma_-)=\frac{R_f}{R_+}-\frac{R_f}{R_-}
\end{equation}
First, we know from \cref{eq:doubleMem2} that the weight is represented by \cref{eq:wei2res0}. This means that the weights are limited in values to :

\begin{itemize}
  \item $w_{max}=R_f\cdot(\sigma_{max}-\sigma_{min})$
  \item $w_{min}=-w_{max}=R_f\cdot(\sigma_{min}-\sigma_{max})$
\end{itemize}

For the rest of this chapter we consider that $R_f$ is set to the middle point of $R_{max}$ and $R_{min}$ meaning that $R_f=\frac{R_{min}+R_{max}}{2}$.

Since we only have one equation (\cref{eq:wei2res0}) for 2 unknowns ($R_+$ and $R_-$), we need to set a second equation. This is done by centering the resistances around $R_f$, this means that \cref{eq:wei2res1} is the second equation, that makes our problem now solvable.

\begin{equation}
  \label{eq:wei2res1}
  R_f=\frac{R_-+R_+}{2}
\end{equation}
\begin{equation}
  \label{eq:wei2res01}
  \begin{cases}
    w=\frac{R_f}{R_+}-\frac{R_f}{R_-}\\
    R_f=\frac{R_-+R_+}{2}
  \end{cases}
\end{equation}

By solving \cref{eq:wei2res01}, we find the \cref{eq:wei2res2} that gives the values for $R_-$ and $R_+$. All the steps for solving the equations can be found in \cref{apsec:wei2res}.

The real value to voltage conversion (\cref{tab:valConv}) does not affect this part of the system as the crossbar array already works in voltages. Replacing the resistor variable in \cref{eq:doubleMem2} by $w$ using \cref{eq:wei2res0} would not change the results.

\begin{equation}
  \label{eq:wei2res2}
  \begin{cases}
    R_+= (w+1-\sqrt{w^2+1})\cdot\frac{R_f}{w}\\
    R_-=2\cdot R_f -R_+
  \end{cases}
\end{equation}

While solving the system (\cref{apsec:wei2res}), we get two potential equations for $R_+$. Graphing them shows that one of them is outside of the memristor's resistance range ($[R_{min},R_{max}]$). Leaving us with only one equation because the other one is physically unreachable.

@@

>>>>

\begin{equation}
  \label{eq:wei2res0}
  \symw=\symR_f\cdot( \symG_+- \symG_-)=\frac{\symR_f}{\symR_+}-\frac{\symR_f}{\symR_-}
\end{equation}
First, we know from \cref{eq:doubleMem2} that the weight is represented by \cref{eq:wei2res0}. This means that the weights are limited in values to :

\begin{itemize}
  \item $\symw_{max}=\symR_f\cdot( \symG_{max}- \symG_{min})$
  \item $\symw_{min}=-\symw_{max}=\symR_f\cdot( \symG_{min}- \symG_{max})$
\end{itemize}

For the rest of this chapter we consider that $\symR_f$ is set to the middle point of $\symR_{max}$ and $\symR_{min}$ meaning that $\symR_f=\frac{\symR_{min}+\symR_{max}}{2}$.

Since we only have one equation (\cref{eq:wei2res0}) for 2 unknowns ($\symR_+$ and $\symR_-$), we need to set a second equation. This is done by centering the resistances around $\symR_f$, this means that \cref{eq:wei2res1} is the second equation, that makes our problem now solvable.

\begin{equation}
  \label{eq:wei2res1}
  \symR_f=\frac{\symR_-+\symR_+}{2}
\end{equation}
\begin{equation}
  \label{eq:wei2res01}
  \begin{cases}
    \symw=\frac{\symR_f}{\symR_+}-\frac{\symR_f}{\symR_-}\\
    \symR_f=\frac{\symR_-+\symR_+}{2}
  \end{cases}
\end{equation}

By solving \cref{eq:wei2res01}, we find the \cref{eq:wei2res2} that gives the values for $\symR_-$ and $\symR_+$. All the steps for solving the equations can be found in \cref{apsec:wei2res}.

The real value to voltage conversion (\cref{tab:valConv}) does not affect this part of the system as the crossbar array already works in voltages. Replacing the resistor variable in \cref{eq:doubleMem2} by \symw using \cref{eq:wei2res0} would not change the results.

\begin{equation}
  \label{eq:wei2res2}
  \begin{cases}
    \symR_+= (\symw+1-\sqrt{\symw^2+1})\cdot\frac{\symR_f}{\symw}\\
    \symR_-=2\cdot \symR_f -\symR_+
  \end{cases}
\end{equation}

While solving the system (\cref{apsec:wei2res}), we get two potential equations for $\symR_+$. Graphing them shows that one of them is outside of the memristor's resistance range ($[\symR_{min},\symR_{max}]$). Leaving us with only one equation because the other one is physically unreachable.

@@

The python implementation is available in \cref{apsec:wei2res}.
